\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{booktabs} 
\usepackage{makecell} 
\usepackage{multirow} 
\usepackage{booktabs}
\usepackage{array}
\usepackage{siunitx} 
\usepackage{pifont}
\usepackage[hidelinks]{hyperref}
\newcommand{\cmark}{\ding{51}}%
\newcommand{\xmark}{\ding{55}}%
\usepackage{tikz}
\usetikzlibrary{patterns.meta,backgrounds, calc, arrows.meta, intersections, patterns, positioning, shapes.misc, fadings, through,decorations.pathreplacing,shapes}
\usepackage{caption}
\usepackage{enumitem}
\usepackage{listofitems}
\usepackage{changepage}
\usepackage{listings}

% Enable or disable reviews
\newif\ifreview
\reviewtrue
%\reviewfalse

\definecolor{amber}{rgb}{1.0, 0.75, 0.0}
\newcounter{ReviewerID}
\readlist*\annotationcolors{blue, red, orange, green, purple, amber, brown, olive}
\newcommand{\newreviewer}[2]{%
        \ifnum \theReviewerID=\annotationcolorslen
        \setcounter{ReviewerID}{0} % cycle through colors
        \fi
        \stepcounter{ReviewerID}%
        \expandafter\edef\csname bootstrap#1\endcsname{%
                \expandafter\def\csname #1\endcsname####1{%
                        \ifreview%
                         {\noexpand\color{\annotationcolors[\theReviewerID]} {\noexpand\bf{\noexpand\fbox{#2}} {\noexpand\it ####1} }}
                        \else%
                         {}% disable annotations
                        \fi%
                }%
        }%
        \csname bootstrap#1\endcsname%
}



\makeatletter
\let\old@lstKV@SwitchCases\lstKV@SwitchCases
\def\lstKV@SwitchCases#1#2#3{}
\makeatother
\usepackage{lstlinebgrd}
\makeatletter
\let\lstKV@SwitchCases\old@lstKV@SwitchCases

\lst@Key{numbers}{none}{%
    \def\lst@PlaceNumber{\lst@linebgrd}%
    \lstKV@SwitchCases{#1}%
    {none:\\%
     left:\def\lst@PlaceNumber{\llap{\normalfont
                \lst@numberstyle{\thelstnumber}\kern\lst@numbersep}\lst@linebgrd}\\%
     right:\def\lst@PlaceNumber{\rlap{\normalfont
                \kern\linewidth \kern\lst@numbersep
                \lst@numberstyle{\thelstnumber}}\lst@linebgrd}%
    }{\PackageError{Listings}{Numbers #1 unknown}\@ehc}}
\makeatother



\newreviewer{yiwen}{yiwen}
\newreviewer{mat}{mat}

\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

\title{Evolution of Rust for Linux: From Unsafe to Safe}

\author{\IEEEauthorblockN{Yiwen Xu}
\IEEEauthorblockA{
\textit{École Polytechnique Fédérale de Lausanne (EPFL)}\\
Lausanne, Switzerland}
% \and
% \IEEEauthorblockN{6\textsuperscript{th} Given Name Surname}
% \IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
% \textit{name of organization (of Aff.)}\\
% City, Country \\
% email address or ORCID}
}

\maketitle

\begin{abstract}
Linux has been the de facto operating system for servers and desktop computers for decades, yet it has been plagued by a number of security vulnerabilities.
A significant cause is that, for performance reasons, the kernel is written in C, a language that is not memory-safe by design. For this reason, the Linux kernel community has been exploring the use of memory-safe languages to improve the security of the kernel.
Rust, as a modern system programming language with security at its core, has gained popularity in the Linux kernel community. Therefore, the Rust for Linux project was initiated to add the fundamental support for Rust development to the Linux kernel, following by many practices on new or old drivers being rewritten in Rust.

Driver development cannot exist independently of the existing kernel subsystems, which poses a significant challenge for the Rust for Linux project: how can idiomatic Rust be integrated elegantly into the existing kernel without disrupting its infrastructure, while still coexisting seamlessly with the remaining C code? 


To help shape the ongoing works of Rust drivers and other Rust--C
systems-level projects, we present a comprehensive study of the evolution
of Rust for Linux, from the initial introduction of Rust support to the
current state-of-the-art design and the disciplined use of \texttt{unsafe}
blocks. 
First, we examine the design choices behind the abstraction layers
that bridge Rust and C. 
Second, we analyze the rationale for \texttt{unsafe}
Rust usage---the only source of residual memory-safety issues---and
trace its evolution across different abstraction layers and drivers in the kernel.
Finally, we distill the practical guidelines and best practices
that have emerged from the Rust for Linux project for future development.

This study is both timely and necessary, as Rust has moved from an
experimental effort to an increasingly established part of kernel development.
Its growing adoption has also intensified interest in rewriting kernel
components and other systems-level software in Rust, especially where existing
C implementations have suffered from recurring memory-safety issues.


\end{abstract}

\begin{IEEEkeywords}
Trusted Execution Environment, Bug Detection, Double Fetch Vulnerabilities
\end{IEEEkeywords}

\section{Introduction}



The Linux kernel has underpinned modern computing infrastructure for over three decades, running on everything from embedded devices and smartphones
to cloud servers and supercomputers. Yet, despite its maturity and widespread deployment, the kernel remains a persistent source of critical security
vulnerabilities. Memory safety errors---buffer overflows, use-after-free, null-pointer dereferences, and out-of-bounds accesses---account for a disproportionate fraction of reported kernel CVEs. 

Empirical analyses of the Linux CVE database consistently show that a majority of high-severity vulnerabilities trace back to memory unsafety~\cite{TODO:cve_study,TODO:linux_memory_bugs}, a fundamental consequence of the kernel being written in C, a language that offers no compile-time guarantees over memory accesses and relies entirely on programmer discipline for correctness. Furthermore, in the LLM era, the number of vulnerabilities reported by LLMs to Linux kernel has increased significantly, making it more difficult to Linux community to patch vulnerabilities in a timely manner~\cite{TODO:llm_vulnerabilities_to_linux}.

Rust~\cite{TODO:rust_book} has emerged as a feasible and principled response to this challenge. As a modern system programming language with memory safety and efficiency as its core design goals, Rust enforces strict ownership, borrowing, and lifetime rules through its type system and borrow checker, statically and dramatically eliminating most classes of memory errors at compile time. Unlike garbage-collected languages, it is crucial to note that Rust achieves these guarantees without a runtime overhead, preserving the low-level control and performance required for kernel development.

Motivated by these properties, the Rust for Linux (RFL) project~\cite{TODO:ojeda2022rust, TODO:rustforlinux} was initiated to introduce Rust as a second implementation language in the Linux kernel. The goal is not to rewrite the existing kernel in Rust, but to allow newly-involved security-critical drivers and kernel subsystems to be developed in idiomatic, safe Rust.
After extensive preparatory work spanning several years, Rust support was merged into the mainline kernel at Linux~v6.1~\cite{TODO:linux_v6.1}. Today, Rust is widely regarded as a successful second language in the kernel community, and its status has continued to mature following the removal of the ``experimental'' label in Linux~v6.19~\cite{TODO:linux_v6.19}.
The project follows a layered architecture: the existing C kernel infrastructure forms the foundation; a thin layer of \texttt{unsafe} Rust bindings bridges Rust to the C~APIs; an elaborately-designed and carefully-reviewed \emph{safe abstraction layer} wraps these bindings and exposes ergonomic, safe Rust interfaces; and Rust drivers are then built on top of this abstraction layer.
This design localizes majority of \texttt{unsafe} blocks, the only source of memory safety issues in Rust code, to the abstraction layer and shields the leaf driver code from direct C interoperability hazards and the associated security risks.

Despite this architecture, Rust's memory safety guarantees are not absolute or unconditional in the kernel context. 
The \texttt{unsafe} keyword marks code regions where the compiler relaxes its safety invariants for five superpowers.
It allows raw pointer dereferences, unsafe function calls (including calls to foreign language interface), mutable static variable access, unsafe casts, and union field operations.
Those superpowers are necessary for the feasible implementation of Rust-based system or programs in terms of performance and functionality, such as interrupt handling, memory management, and low-level device drivers.
Especially, in Rust for Linux, \texttt{unsafe} blocks are an unavoidable necessity: the fundamental mismatch between Rust's strict ownership model and C's unconstrainted type system means that no safe abstraction can be built without an \texttt{unsafe} foundation.
The project therefore adopts a discipline: \texttt{unsafe} code should be confined to the abstraction layers with rigorous documentation and review process, so that driver code above can remain \texttt{unsafe}-free.

Prior research has examined Rust security mainly for the broader Rust ecosystem and concentrates on automated detection of \texttt{unsafe} bugs.
RUDRA~\cite{TODO:bae2021rudra} performs inter-procedural dataflow analysis to find unsound patterns in unsafe Rust APIs, targeting violations of higher-order safety invariants and incorrect \texttt{Send}/\texttt{Sync} implementations commonly found in published crates.
SafeDrop~\cite{TODO:safedrop} targets use-after-free and double-free vulnerabilities by tracking ownership transfer and drop semantics through a path-sensitive analysis of Rust's Mid-level Intermediate Representation(MIR), making it effective at catching lifetime-related misuse.
TypePulse~\cite{TODO:typepulse} takes a type-confusion-centric approach, identifying bugs that arise from unsound type casts in \texttt{unsafe} code, particular relevance at foreign language interface (FFI) boundaries where Rust and C types are coerced into one another.
While all three tools target \texttt{unsafe}-related memory bugs, they differ in their underlying representations, targeted vulnerability classes, and scope: RUDRA and SafeDrop are designed for the Rust crate ecosystem, whereas TypePulse is specifically oriented toward cross-language type safety at FFI boundaries.

Another line of Rust research takes an empirical perspective on how \texttt{unsafe} code is actually adopted across the Rust ecosystem.
Evans et al.~\cite{TODO:evans2020rust} conducted one of the first large-scale studies of \texttt{unsafe} usage across crates.io, finding that although fewer than 30\% of libraries invoke the \texttt{unsafe} keyword directly, the majority cannot be fully statically verified by the compiler because unsafeness propagates transitively through dependency call chains.
Astrauskas et al.~\cite{TODO:astrauskas2020programmers} subsequently evaluated the so-called \emph{Rust hypothesis}, which holds that programmers use \texttt{unsafe} sparingly, keep it easy to review, and hide it behind safe abstractions. They tested this hypothesis by querying a large corpus of crates with their QRATES tool over both source code and MIR. They identified six recurring motivations for \texttt{unsafe} code and reported that the hypothesis holds only partially.
Cui et al.~\cite{TODO:cui2024achilles} approached the problem from the perspective of safety requirements rather than usage frequency. By analyzing the standard library's \texttt{unsafe} API documentation, they distilled 19 safety properties, labeled 416 \texttt{unsafe} APIs, and cross-referenced them against historical Rust CVEs and crates.io usage, revealing that the documented safety requirements are frequently inconsistent and incomplete. It poses the challenge to Rust programmers to understand and fulfill the safety requirements accordingly.

Different from these studies all characterizing \texttt{unsafe} usage or obligations of Rust-based codebases across the open-source crate ecosystem at large, Li et al.~\cite{TODO:li2024empirical} conducted the first large-scale qualitative study of Rust's integration into the C-written Linux kernel.
It examines the Rust for Linux project through developer mailing lists, patch discussions, and issue reports, documenting sources of success, dissatisfaction, and compromise in the project's early stages.
While this study provides valuable sociological and high-level architectural insight, it is necessarily broad: it was conducted before the project had reached sufficient maturity to include complex, production-quality driver examples, and it does not systematically analyze the safety properties and  changes of abstraction layers, nor does it feature how \texttt{unsafe} is utilized and distributed within it.

In this paper, we present the systematic and in-depth study of the evolution of the non-experimental Rust for Linux, with a focus on the design and safety of its abstraction layers and the disciplined use of \texttt{unsafe} Rust in real kernel code. Our study shows that \texttt{unsafe} usage is unevenly distributed across Rust-for-Linux and the abstraction layer continues evolving as the new drivers are introduced and the project matures, instead of directly mirroring the existing C codebase.

We trace the Rust integration on how the abstraction layers have been designed and refined over successive kernel releases and analyze the pros and cons it brings compared to the existing C codebase. 
Further, we systematically summarize the rationale behind \texttt{unsafe} usage, and distill the practical guidelines and best practices that have emerged from the project for future Rust kernel development.

\medskip
\noindent\textbf{Contributions.} This paper makes the following contributions:

\begin{itemize}[leftmargin=*]
    \item \textbf{Architectural analysis.} We provide a comprehensive
    longitudinal analysis of the Rust for Linux abstraction layer design,
    tracing its evolution from the initial Rust support in Linux~v6.1 to the
    current state of the art, and characterizing the key primitives that
    govern the Rust-C interoperability.

    \item \textbf{Unsafe characterization.} We systematically study the
    distribution, rationale, and evolution of \texttt{unsafe} blocks across
    abstraction layers and drivers in the Rust for Linux codebase, revealing
    consistent patterns in how and why \texttt{unsafe} is utilized and how these
    patterns have changed as the project matures.

    \item \textbf{Design guidelines.} We distill a set of practical design
    principles and best practices for writing safe Rust kernel code, grounded
    in the real experience of the Rust for Linux project and intended to
    guide future development of other Rust-C system-level projects.
\end{itemize}

\input{sections/background}

\input{sections/design}

\input{sections/unsafe}

% \input{sections/practice}

\section{Conclusion}

This paper presented a systematic study of the evolution of Rust for Linux,
examining how the project has matured from its initial experimental
infrastructure in Linux~v6.1 to a non-experimental, production-quality
second language in the kernel.
Our analysis of the abstraction layer architecture reveals that the design
has evolved incrementally but deliberately. The \texttt{unsafe} code is
consistently confined to the Rust-C binding and abstraction layer, while
driver code built on top remains largely free of \texttt{unsafe} blocks.

Our characterization of \texttt{unsafe} usage further shows that the
dominant motivations, including FFI calls, raw pointer manipulation, and shared
mutable state, are inherent to the C~interface rather than to poor Rust
practice, underscoring that \texttt{unsafe} in the kernel context is
a necessary and contained cost rather than a sign of unsound design.

Different from the insights from Rust crate ecosystem, any project that integrates Rust into a large, pre-existing C codebase have to face the same fundamental tension between Rust's ownership model and C's unconstrained type system during their co-existence.
Importantly, our findings can help practitioners establish robust practices for Rust-involving system-level software projects, supporting their transition from promising prototypes to reliable infrastructure.

\bibliographystyle{IEEEtran}
\bibliography{IEEEexample}


\end{document}
