\section{Design of Rust for Linux}

\subsection{Design Principles}

\textbf{Orientation:} Rust for Linux introduces Rust as a second implementation language choice alongside C, not as a replacement for it.
A complete kernel rewrite is neither realistic nor desirable, as the existing C codebase embodies decades of carefully tuned logic and a wholesale migration would introduce far greater risk than it eliminates.
Instead, the project targets newly developed drivers, kernel modules and necessary subsystems where Rust’s compile-time memory-safety guarantees deliver the greatest benefit. This co-existence strategy is not cost-free. Inherent incompatibilities persist between Rust’s strict ownership model and C’s unconstrained pointer semantics, requiring significant manual effort for integration and API redesign or wrapping on each C subsystem basis. Shared kernel mechanisms such as locks, RCU, and workqueues cannot be reused directly from C and must each be re-expressed carefully through a Rust abstraction that encodes the relevant safety invariants in the type system.

\textbf{Evolution and Core Design:} To contain this complexity, the project adopts an incremental upstreaming strategy by first merging the core Rust infrastructure (toolchain, build system, C bindings, and thin glue code), then upstreaming a small number of simple reference drivers to validate the approach, and finally gradually expanding and improving the abstraction to more subsystems over time. 
The architecture enforces a strict three-layer design illustrated in Figure~\ref{fig:rfl_arch}.
At the bottom, \emph{bindings} provide raw Rust declarations of the C types,
functions, and constants from \texttt{include/}, auto-generated by the \texttt{bindgen} tool from kernel headers.
Constructs that \texttt{bindgen} cannot handle, such as \texttt{inline} functions or complex macros, are exposed through small C wrapper functions
placed in \texttt{rust/helpers/}.
Above the bindings sits the \emph{safe abstraction layer} in
\texttt{rust/kernel/}, wrapping the unsafe binding access into an
as-safe-as-possible Rust API.
At the top, leaf modules such as drivers and filesystems are built
exclusively against this safe layer.
Direct access from any leaf module to the raw bindings crate is explicitly
\textbf{forbidden} so that all crossings of the Rust-C boundary remain
concentrated at a single, reviewable boundary.
The only external Rust dependencies are the upstream \texttt{core} crate for
language primitives and \texttt{alloc} for heap allocation, where
\texttt{alloc} was maintained as an in-tree fork until Linux~v6.10 to add
fallible (\texttt{try\_*}) allocation variants returning \texttt{Result}
rather than panicking, a hard requirement for kernel code.

\input{figures/overview.tex}

The central design principle of this architecture is \emph{encapsulation},
meaning that every direct interaction with C kernel APIs must be confined to
carefully reviewed and documented abstractions.
Under this invariant, users of the abstraction layer cannot introduce
undefined behavior (UB) provided two conditions hold.
First, the abstractions must be \emph{sound} in the sense that their
\texttt{unsafe} implementations correctly uphold the safety invariants they
promise to callers.
Second, any \texttt{unsafe} block in user code must respect the documented
safety contract of the operation it invokes, and any \texttt{unsafe impl}
must correctly fulfill the safety contract of the trait being implemented.
This design concentrates the burden of low-level reasoning at a single layer
so that abstraction authors bear the safety responsibility while driver code
above can remain free of \texttt{unsafe} blocks entirely.
Beyond safety, abstractions are also expected to be \emph{ergonomic} in the
sense that C resource acquisition and release patterns are mapped to Rust
constructors and \texttt{Drop} destructors, integer error codes become
\texttt{Result} types, and kernel object lifecycles are encoded as Rust
lifetime annotations, making driver code concise, idiomatic, and readable.
Abstraction coverage is intentionally incomplete at any given release and
expands incrementally as new drivers are upstreamed, which means leaf modules
must never work around a missing abstraction by reaching directly into the
bindings layer.

\subsection{Primitives of the Safe Abstraction Layer}

The safe abstraction layer in \texttt{rust/kernel/} is not built from ad-hoc wrappers but from a
small set of reusable primitives that each address a specific mismatch between Rust's type system
and C's memory model.
Every higher-level abstraction---locks, device registrations, file operations, bus drivers---is
ultimately composed from these building blocks.
Five key primitives are described below.

\textbf{Opaque\textless T\textgreater.}
The most fundamental primitive is \texttt{Opaque<T>}, defined in \texttt{rust/kernel/types.rs}.
Whenever a Rust struct embeds a C struct that will be modified or initialized by the kernel's own
C code, wrapping it in a plain Rust field would violate several of Rust's core assumptions: the
memory may not be fully initialized, C may mutate it through an aliased pointer while a shared
Rust reference exists, and the value must not be moved after it has been registered with the
kernel.
\texttt{Opaque<T>} resolves all three issues at once.
The \texttt{UnsafeCell} removes the immutability guarantee normally assumed for shared references,
\texttt{MaybeUninit} suppresses the initialized-value assumption, and \texttt{PhantomPinned} makes
the type \texttt{!Unpin} so the compiler refuses to move it after pinning:

\begin{lstlisting}[language=Rust,columns=fullflexible]
#[repr(transparent)]
pub struct Opaque<T> {
    value: UnsafeCell<MaybeUninit<T>>,
    _pin: PhantomPinned,
}
\end{lstlisting}

\noindent
The only way to access the inner value is through the raw pointer returned by \texttt{Opaque::get},
which forces every read or write to be wrapped in an \texttt{unsafe} block with an explicit safety
comment.
Two initialization helpers accompany the type: \texttt{Opaque::ffi\_init} and
\texttt{Opaque::try\_ffi\_init}, which accept a closure receiving a raw \texttt{*mut T} and
integrate with the \texttt{pin\_init!} macro.
The pattern permeates the abstraction layer.
As a concrete example, the generic \texttt{Lock<T,~B>} type in \texttt{rust/kernel/sync/lock.rs}
stores the C lock state as \texttt{state:~Opaque<B::State>} and initializes it as:

\begin{lstlisting}[language=Rust,columns=fullflexible]
state <- Opaque::ffi_init(|slot| unsafe {
    B::init(slot, name.as_char_ptr(), key.as_ptr())
}),
\end{lstlisting}

\noindent
The closure forwards the raw slot pointer directly to the C \texttt{mutex\_init} or
\texttt{spin\_lock\_init} call, while the surrounding \texttt{pin\_init!} macro guarantees that
the containing struct is never moved afterward.
The same pattern applies to \texttt{work\_struct} in \texttt{workqueue.rs}, \texttt{pwm\_chip} in
\texttt{pwm.rs}, and driver-registration blobs in \texttt{driver.rs}.

\textbf{ForeignOwnable.}
The second key primitive governs the transfer of heap-allocated Rust objects across the FFI
boundary.
Many kernel subsystems store driver-private state as a type-erased \texttt{void~*} inside a
C struct---for example, \texttt{dev->driver\_data}, \texttt{file->private\_data}, or
\texttt{policy->driver\_data}.
Without a disciplined convention this pattern leads to double-frees or use-after-free bugs: the
Rust allocator and the C side would each believe they own the same allocation.
The \texttt{ForeignOwnable} trait, defined in \texttt{rust/kernel/types.rs}, makes the ownership
protocol explicit and compiler-checked:

\begin{lstlisting}[language=Rust,columns=fullflexible]
pub unsafe trait ForeignOwnable: Sized {
    fn into_foreign(self) -> *mut c_void;
    unsafe fn from_foreign(ptr: *mut c_void) -> Self;
    unsafe fn borrow<'a>(ptr: *mut c_void)
        -> Self::Borrowed<'a>;
    unsafe fn borrow_mut<'a>(ptr: *mut c_void)
        -> Self::BorrowedMut<'a>;
}
\end{lstlisting}

\noindent
Calling \texttt{into\_foreign} transfers ownership from Rust to C by consuming the value and
returning a raw pointer; the borrow checker will refuse any further use of the original binding.
\texttt{from\_foreign} performs the reverse transfer and must be called at most once per pointer.
\texttt{borrow} and \texttt{borrow\_mut} allow temporary access while C still holds the pointer,
and the lifetime parameter \texttt{'a} prevents any reference from outliving that pointer.
Implementations are provided for \texttt{Box<T>}, \texttt{Pin<Box<T>>}, and \texttt{Arc<T>}.
The platform bus driver in \texttt{rust/kernel/platform.rs} demonstrates the idiom across the
probe/remove lifecycle:

\begin{lstlisting}[language=Rust,columns=fullflexible]
// probe: allocate driver state and hand ownership to C
pdev.as_ref().set_drvdata(data)?;

// remove: borrow without reclaiming ownership
let data = unsafe { pdev.as_ref().drvdata_borrow::<T>() };
T::unbind(pdev, data);
\end{lstlisting}

\noindent
Internally \texttt{set\_drvdata} calls \texttt{KBox::pin\_init} followed by
\texttt{into\_foreign} and \texttt{dev\_set\_drvdata}, while the device resource manager in
\texttt{devres.rs} closes the lifecycle by registering a C cleanup callback that calls
\texttt{from\_foreign} and \texttt{drop} when the device is unbound, ensuring the allocation is
freed exactly once.

\textbf{Callback Bridge.}
A third recurring challenge is connecting the kernel's C-side vtables---\texttt{file\_operations},
\texttt{platform\_driver}, \texttt{pwm\_ops}, and similar structs of function pointers---to safe
Rust trait implementations.
Rust closures cannot be placed directly into C function-pointer fields because they carry a fat
pointer (code + data) and may have a destructor, neither of which C understands.
The solution used throughout the abstraction layer is a thin \texttt{extern "C"} adapter function,
which the compiler emits as an ordinary C-ABI symbol, that recovers the Rust context pointer and
delegates to the trait method.
The misc-device abstraction in \texttt{rust/kernel/miscdevice.rs} illustrates the full
open/release cycle:

\begin{lstlisting}[language=Rust,columns=fullflexible]
extern "C" fn open_callback(...) -> c_int {
    let ptr = match T::open(file, misc) {
        Ok(ptr) => ptr,
        Err(err) => return err.to_errno(),
    };
    // Transfer ownership to C via ForeignOwnable
    unsafe { (*raw_file).private_data = ptr.into_foreign() };
    0
}

extern "C" fn release_callback(...) -> c_int {
    let private = unsafe { (*file).private_data };
    // Reclaim ownership and drop
    let ptr = unsafe {
        <T::Ptr as ForeignOwnable>::from_foreign(private)
    };
    T::release(ptr, ...);
    0
}
\end{lstlisting}

\noindent
The adapter functions are zero-cost: because the Rust trait method is known at compile time, the
compiler inlines or calls it directly without any virtual dispatch overhead.
The \texttt{\#[vtable]} procedural macro, defined in \texttt{rust/macros/vtable.rs}, further
refines this pattern by generating a \texttt{HAS\_\textit{METHOD}} boolean constant for each
optional operation in the trait.
When a driver does not implement an optional method, the corresponding function-pointer field in
the C vtable is set to \texttt{NULL} at compile time, eliminating the runtime check entirely.
The same bridge pattern is used for the platform, PCI, USB, and I2C bus drivers, PWM chip
operations, and debugfs file operations.

\textbf{PinInit / \texttt{pin\_init!}.}
A fundamental problem unique to kernel Rust is that many C objects---mutexes, spinlocks,
work-structs, completion structs---must be \emph{initialized in place} at a fixed memory address,
because the kernel registers that very address with internal data structures such as lock-dependency
tracking or watchdog lists.
Rust's ordinary constructor syntax first builds a value on the stack and then \emph{moves} it to
its final location, which would silently corrupt any address-sensitive C object.
The \texttt{pin-init} crate (in-tree at \texttt{rust/pin-init/}) solves this by representing
initialization as a \emph{closure over a raw pointer}: instead of returning a value, an initializer
receives a \texttt{*mut Self} pointing at the final allocation and writes directly into it, so no
move ever occurs.
The \texttt{\#[pin\_data]} attribute on a struct, combined with \texttt{\#[pin]} on individual
fields, teaches the \texttt{pin\_init!} macro which fields require in-place initialization and
which can be moved normally:

\begin{lstlisting}[language=Rust,columns=fullflexible]
#[pin_data]
struct SampleDriver {
    #[pin]
    state: Mutex<u32>,       // must stay pinned: wraps C mutex
    name: CString,           // movable Rust value
}

impl platform::Driver for SampleDriver {
    fn probe(pdev: &platform::Device<Core>, _: Option<&Self::IdInfo>)
        -> impl PinInit<Self, Error>
    {
        try_pin_init!(Self {
            state <- new_mutex!(0),
            name:  CString::try_from_fmt(fmt!("sample"))?,
        })
    }
}
\end{lstlisting}

\noindent
The \texttt{<-} arrow syntax invokes an in-place initializer for pinned fields; the \texttt{:}
syntax moves ordinary values.
The macro expands to a closure that calls each field's initializer in sequence against the target
pointer, rolling back already-initialized fields on error so that partial initialization never
leaks.
\texttt{KBox::pin\_init} and \texttt{Arc::pin\_init} allocate heap memory and immediately pass the
pointer to the initializer, so the constructed object never touches the stack at all.
This mechanism is the foundation on which \texttt{Opaque::ffi\_init}, every lock type, and every
driver probe function in RFL is built.

\textbf{ARef\textless T\textgreater{} and AlwaysRefCounted.}
The final primitive addresses a mismatch in reference-counting models.
Many core kernel objects---\texttt{struct device}, \texttt{struct net\_device},
\texttt{struct task\_struct}---carry their own C-managed reference count, incremented and
decremented by functions such as \texttt{get\_device} / \texttt{put\_device}.
Wrapping such an object in Rust's \texttt{Arc<T>} would introduce a \emph{second} refcount
layered on top of the C one, causing the allocation to be freed by the Rust side when its own
count reaches zero while the C side still holds live references---a use-after-free bug.
The \texttt{ARef<T>} smart pointer, defined in \texttt{rust/kernel/sync/aref.rs}, solves this by
delegating entirely to the object's internal refcount through the \texttt{AlwaysRefCounted} trait:

\begin{lstlisting}[language=Rust,columns=fullflexible]
pub unsafe trait AlwaysRefCounted {
    fn inc_ref(&self);
    unsafe fn dec_ref(obj: NonNull<Self>);
}

pub struct ARef<T: AlwaysRefCounted> {
    ptr: NonNull<T>,
    _p: PhantomData<T>,
}
\end{lstlisting}

\noindent
Cloning an \texttt{ARef<T>} calls \texttt{inc\_ref}; dropping it calls \texttt{dec\_ref}---both
of which delegate to the corresponding C functions.
No additional allocation is involved; \texttt{ARef} is layout-identical to a single raw pointer.
In practice, every driver that needs to hold a device handle past the end of the probe callback
stores an \texttt{ARef<platform::Device>} (or \texttt{ARef<pci::Device>}, etc.) in its private
data struct.
The platform sample in \texttt{samples/rust/rust\_driver\_platform.rs} illustrates this directly:

\begin{lstlisting}[language=Rust,columns=fullflexible]
struct SampleDriver {
    pdev: ARef<platform::Device>,
}

// In probe:
Ok(Self { pdev: pdev.into() })
\end{lstlisting}

\noindent
The \texttt{pdev.into()} call increments the kernel refcount and stores the result; when
\texttt{SampleDriver} is eventually dropped, \texttt{ARef}'s \texttt{Drop} implementation
decrements it.
The same pattern appears in every sample driver that retains a device reference:
\texttt{rust\_soc.rs}, \texttt{rust\_dma.rs}, \texttt{rust\_driver\_pci.rs},
\texttt{rust\_misc\_device.rs}, and \texttt{rust\_driver\_usb.rs}.

\subsection{Examples of Binder Drivers}

\begin{figure*}[t]
\centering
\begin{tikzpicture}[
  %% node styles
  vfsnode/.style  = {draw=gray!60,       fill=gray!12,    rounded corners=3pt,
                     minimum width=10.4cm, minimum height=0.80cm,
                     align=center, font=\small},
  cnode/.style    = {draw=orange!70!black,fill=orange!10,  rounded corners=3pt,
                     minimum width=4.6cm, minimum height=0.75cm,
                     align=center, font=\small},
  rnode/.style    = {draw=blue!60!black,  fill=blue!8,     rounded corners=3pt,
                     minimum width=4.6cm, minimum height=0.75cm,
                     align=center, font=\small},
  typenode/.style = {draw=blue!40!black,  fill=blue!4,     rounded corners=2pt,
                     minimum width=4.2cm, minimum height=0.65cm,
                     align=center, font=\scriptsize},
  badnode/.style  = {draw=red!55!black,   fill=red!6,      rounded corners=2pt,
                     minimum width=4.2cm, minimum height=0.65cm,
                     align=center, font=\scriptsize},
  annot/.style    = {font=\scriptsize\itshape, align=center},
  %% arrow styles
  main/.style  = {-Stealth, thick},
  saf/.style   = {-Stealth, thick, blue!55!black},
  uns/.style   = {-Stealth, thick, orange!55!black},
  own/.style   = {-Stealth, thick, teal!60!black},
]

%% -- VFS / binderfs (top, full width) ----------------------------------
\node[vfsnode] (vfs) at (0, 0)
  {\textbf{VFS / binderfs (shared C layer)}\\[-2pt]
   {\scriptsize\texttt{inode->i\_fop = \&\{c,rust\}\_binder\_fops}\quad
    \texttt{open / ioctl / mmap / release / poll / flush}}};

%% -- Column headers -----------------------------------------------------
\node[annot, orange!70!black] at (-2.8, -1.15) {\textbf{C Binder} (\texttt{binder.c})};
\node[annot, blue!70!black]   at ( 2.8, -1.15) {\textbf{Rust Binder} (\texttt{drivers/android/binder/})};

%% -- C column -----------------------------------------------------------
\node[cnode] (calloc) at (-2.8, -1.90)
  {\texttt{kzalloc} $\rightarrow$ \texttt{binder\_proc *}};

\node[badnode] (crawptr) at (-2.8, -2.85)
  {raw \texttt{*binder\_proc} in \texttt{filp->private\_data}\\
   \textcolor{red!70!black}{\textit{no lifetime, no compiler check}}};

\node[badnode] (crefcnt) at (-2.8, -3.80)
  {manual \texttt{refcount\_inc} / \texttt{refcount\_dec}\\
   \textcolor{red!70!black}{\textit{convention-enforced pairing}}};

\node[badnode] (clist) at (-2.8, -4.75)
  {\texttt{work\_struct} may be inserted into\\
   any list, or twice $\rightarrow$ silent corruption\\
   \textcolor{red!70!black}{\textit{no structural guard}}};

%% -- Rust column --------------------------------------------------------
\node[rnode] (rmain) at (2.8, -1.90)
  {\textbf{\texttt{rust\_binder\_main.rs}}\\[-1pt]
   {\scriptsize registers \texttt{rust\_binder\_fops}; module init/exit}};

\node[typenode] (rctx) at (2.8, -2.85)
  {\texttt{context.rs}\quad\textcolor{teal!70!black}{\texttt{Arc<Context>}}\\[-1pt]
   entry via \texttt{ArcBorrow<'\_, Context>}};

\node[typenode] (rproc) at (2.8, -3.80)
  {\texttt{process.rs}\quad\textcolor{teal!70!black}{\texttt{Arc<Process>}}\\[-1pt]
   \texttt{ForeignOwnable} $\leftrightarrow$ \texttt{file->private\_data}};

\node[typenode] (rthread) at (2.8, -4.75)
  {\texttt{thread.rs}\quad\textcolor{teal!70!black}{\texttt{Arc<Thread>}}\\[-1pt]
   entry via \texttt{ArcBorrow<'\_, Process>}};

\node[typenode] (rtx) at (2.8, -5.70)
  {\texttt{transaction.rs}\\[-1pt]
   \texttt{Work<T,\,ID>} $+$ \texttt{ListArc<T,\,ID>}:
   \textcolor{blue!70!black}{\textit{at most one list slot per ID}}};

%% -- Vertical flow arrows -----------------------------------------------
\draw[uns] (vfs.south -| calloc) -- (calloc.north);
\draw[uns] (calloc.south) -- (crawptr.north);
\draw[uns] (crawptr.south) -- (crefcnt.north);
\draw[uns] (crefcnt.south) -- (clist.north);

\draw[saf] (vfs.south -| rmain) -- (rmain.north);
\draw[own] (rmain.south) -- (rctx.north);
\draw[own] (rctx.south)  -- (rproc.north);
\draw[own] (rproc.south) -- (rthread.north);
\draw[own] (rthread.south) -- (rtx.north);

%% -- Column divider -----------------------------------------------------
\draw[gray!40, dashed, line width=0.6pt] (0, -1.50) -- (0, -6.15);

%% -- Zone shading -------------------------------------------------------
\begin{scope}[on background layer]
  \fill[orange!4, rounded corners=4pt] (-5.4,-1.50) rectangle (-0.05,-6.15);
  \draw[orange!35, rounded corners=4pt, dashed, line width=0.6pt]
                                       (-5.4,-1.50) rectangle (-0.05,-6.15);
  \fill[blue!3,   rounded corners=4pt] ( 0.05,-1.50) rectangle ( 5.4,-6.15);
  \draw[blue!30,  rounded corners=4pt, dashed, line width=0.6pt]
                                       ( 0.05,-1.50) rectangle ( 5.4,-6.15);
\end{scope}

\end{tikzpicture}
\caption{%
  Structural comparison of the C and Rust Android Binder drivers.
  Both register identical \texttt{file\_operations} with the shared \texttt{binderfs} layer (top),
  so the VFS dispatch path is unchanged.
  The C driver (left) uses raw pointers, manually paired reference counts, and unguarded
  list insertions.
  The Rust driver (right) replaces these with \texttt{Arc}-based ownership tracked by the
  compiler, \texttt{ForeignOwnable} for safe \texttt{file->private\_data} transfer,
  lifetime-annotated \texttt{ArcBorrow} for entry-point parameters, and
  \texttt{ListArc<T,\,ID>} to enforce the single-list-slot invariant at compile time.}
\label{fig:binder_design}
\end{figure*}

Android Binder is the core IPC mechanism of Android, enabling cross-process method calls between applications and system services.
The Rust Binder driver (\texttt{drivers/android/binder/}) reimplements the same functionality in modular Rust, with the main entry point in \texttt{rust\_binder\_main.rs} and logic split across dedicated modules: \texttt{process.rs}, \texttt{context.rs}, \texttt{transaction.rs}, \texttt{thread.rs}, and others.
This mirrors the structure of the C driver (\texttt{binder.c}), but the Rust version enforces modularity at the language level through the module system and explicit visibility rules, whereas the C version relies on coding conventions.
Despite being a complete reimplementation, the Rust and C drivers are architecturally aligned at the top level: both register the same \texttt{file\_operations} vtable (\texttt{open}, \texttt{release}, \texttt{flush}, \texttt{ioctl}, \texttt{mmap}, \texttt{poll}) with the kernel's \texttt{fops} dispatch table, and both share the same binderfs layer.
This co-existence is made possible by an elegant abstraction boundary: the Rust driver exports a static \texttt{rust\_binder\_fops} object, and \texttt{rust\_binderfs.c} assigns \texttt{inode->i\_fop = \&rust\_binder\_fops}, so the VFS dispatches identically to Rust or C binder depending solely on which \texttt{fops} pointer is installed---no visible difference exists to callers above.

The most significant differences between the Rust and C drivers stem from RAII (Resource Acquisition Is Initialization) and ownership encoding.
In the C \texttt{binder\_open}, the process struct is allocated with \texttt{kzalloc}, reference counts for the device and task are incremented manually (\texttt{refcount\_inc}, \texttt{get\_task\_struct}), and the pointer is stored raw in \texttt{filp->private\_data} with no compiler-enforced lifetime.
The corresponding Rust \texttt{rust\_binder\_open} instead calls \texttt{Process::open}, wraps the result in an \texttt{Arc<Process>}, and transfers ownership across the FFI boundary by calling \texttt{process.into\_foreign()}, which stores the pointer in \texttt{file->private\_data} in a type-erased but ownership-tracked form.
In \texttt{rust\_binder\_release}, the reverse call \texttt{Arc::<Process>::from\_foreign((*file).private\_data)} reclaims that ownership, and the \texttt{Arc} destructor handles deallocation automatically when the last reference is dropped---eliminating the manual \texttt{kfree} and matching \texttt{refcount\_dec} that the C driver must pair by convention.

Rust's type system provides additional memory-safety guarantees that the C binder cannot express.
A first example is the \texttt{Work<T, ID>} struct in \texttt{rust/kernel/workqueue.rs}, which wraps the C \texttt{work\_struct} for deferred execution:

\begin{lstlisting}[language=Rust,columns=fullflexible]
#[pin_data]
#[repr(transparent)]
pub struct Work<T: ?Sized, const ID: u64 = 0> {
    #[pin]
    work: Opaque<bindings::work_struct>,
    _inner: PhantomData<T>,
}
\end{lstlisting}

\noindent
The \texttt{PhantomData<T>} field stores no actual \texttt{T} value at runtime, but tells the compiler that \texttt{Work<T>} logically owns or depends on values of type \texttt{T}.
This propagates \texttt{T}'s variance, \texttt{Send}/\texttt{Sync} constraints, and drop-check semantics to \texttt{Work<T>}, ensuring the compiler refuses to enqueue a work item that outlives the data it references---a class of use-after-free bug that is entirely invisible to C.
The same principle applies in \texttt{ARef<T>}, where \texttt{PhantomData<T>} informs the borrow checker that the smart pointer conceptually holds a \texttt{T}, so any reference obtained through it cannot outlive the pointer itself.
A second example is \texttt{ListArc<T,~ID>} in \texttt{rust/kernel/list/arc.rs}, a wrapper around \texttt{Arc} that enforces the invariant that at most one \texttt{ListArc} per \texttt{ID} exists for any given object at any point in time.
In C binder, the same \texttt{work\_struct} can be accidentally inserted into two different lists or enqueued twice, corrupting list pointers silently at runtime.
The Rust binder's \texttt{Node}, \texttt{NodeDeath}, and \texttt{DTRWrap} types all carry their linked-list membership through \texttt{ListArc}, so any second insertion attempt is rejected at compile time---no runtime guard is needed.
A third example is \texttt{ArcBorrow<'\_, T>}, a non-owning, lifetime-carrying borrow of an \texttt{Arc}-managed object used throughout the binder for all major entry points:

\begin{lstlisting}[language=Rust,columns=fullflexible]
pub(crate) fn open(ctx: ArcBorrow<'_, Context>, ...) -> Result<Arc<Process>>
pub(crate) fn ioctl(this: ArcBorrow<'_, Process>, ...) -> Result
pub(crate) fn mmap(this: ArcBorrow<'_, Process>, ...) -> Result
\end{lstlisting}

\noindent
In C binder the corresponding functions accept a bare \texttt{struct binder\_proc~*} with no information about how long the pointer remains valid.
\texttt{ArcBorrow<'\_, T>} encodes the liveness constraint directly in the type: the lifetime \texttt{'\_} prevents any callee from storing the borrow beyond the scope of the caller's \texttt{Arc}, making use-after-free for these paths a compile-time error rather than a runtime risk.
Together, \texttt{PhantomData}, \texttt{ListArc}, and \texttt{ArcBorrow} demonstrate how Rust binder achieves the same runtime structure as C binder while encoding correctness invariants that the C type system cannot represent.